---
title: "STAT 545 Class Meeting 08"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```


## Today's Lessons

Today's topic is _one-table verbs with `dplyr`_. We've already seen `select`, `filter`, and `arrange`. 

- mutate
```{r}
gapminder %>% 
  mutate(gdp=gdpPercap*pop,
         gdpBill= round(gdp/10^9,2))
```

- grouped summarize
```{r}
gapminder %>% 
  transmute(gdp=gdpPercap*pop,
         gdpBill= round(gdp/10^9,2))
```
```{r}
#replace h to H
if_else(letters=="h","H", letters)
gapminder %>% 
  mutate(lifeExp = if_else(country=="Canada"& year==1952, 70, lifeExp)) %>% 
  filter(country=="Canada")
```

- grouped mutate

...and their corresponding theoretical function types. 

## Resources

Concepts from today's class are closely mirrored by the following resources.

- Jenny's tutorial on [all the one-table verbs](http://stat545.com/block010_dplyr-end-single-table.html)

Other resources:

- Like learning from a textbook? Check out all of [r4ds: transform](http://r4ds.had.co.nz/transform.html).
- The [intro to `dplyr` vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html) is also a great resource. 

Resources for specific concepts:

- To learn more about window functions and how dplyr handles them, see the [window-functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html) vignette for the `dplyr` package. 

## Participation

To get participation points for today, we'll be filling out the cm008-exercise.Rmd file ([Rmd](https://github.com/STAT545-UBC/Classroom/blob/master/notes/cm008-exercise.Rmd), or [from html](http://stat545.com/Classroom/notes/cm008-exercise.nb.html)), and adding it to your participation repo. 

Let's take time to get set up:

1. Download the file as `.Rmd`. 
2. Move the file to your local participation repo.
3. Open the `.Rmd` file, and change the output type to `github_document`.
4. _Optional, but recommended_: Stage and commit the Rmd file (you can do this through RStudio).
    - You can still get participation marks by uploading the final files to GitHub.
